### 2021-09-05 report/10-introduction.tex
Add README, .gitignore, & .tex files for report

\chapter{Introduction} \label{chapter:Introduction}
\section{Theory}
Here the basic principles behind the project should be explained and any relevant theory given. The present work should be put into perspective in relation to previous investigations and references to these should be included.
\section{Sample text}
\blindtext
\blindtext
\subsection{Sample text in detail}
\blindtext
\blindtext
\blindtext
\section{Another section with lots of sample text}
\blindtext
\blindtext
\blindtext
\blindtext
\blindtext
\blindtext
\blindtext
\blindtext
\blindtext
\blindtext
\blindtext
\blindtext
\blindtext
\blindtext
\blindtext
\blindtext
\blindtext
\blindtext
\blindtext