### 2021-09-05 report/report.tex

\documentclass[sotoncolour]{uosthesis}  % Use the Thesis Style with custom link colour
\graphicspath{{../figures/}}            % Location of your graphics files
\usepackage[round]{natbib}              % Use Natbib style for the refs.
\usepackage{bibentry}                   % Use bibentry for prepublished works
\nobibliography*                        % Use bibentry for prepublished works
\usepackage{attrib}                     % Use the attrib package for quotations
\hypersetup{colorlinks=true}            % Set to false for black/white printing
\hypersetup{pdfstartpage=1}             % Set first page to skip copyright declaration

\usepackage{blindtext}                  % TODO: removeme

%% ----------------------------------------------------------------
%% --------------------THESIS/DOC INFORMATION ---------------------
%\faculty     {Faculty of Science}
%\FACULTY     {\MakeUppercase{\facname}}
\department  {School of Physics and Astronomy}
\DEPARTMENT  {\MakeUppercase{\deptname}}
%\group       {[Group name]}
%\GROUP       {\MakeUppercase{\groupname}}
\title       {BSc Project Report}
\modulename  {PH4111}
\authors     {180014855}
\date        {\today}

\begin{document}
%% ------------------ FRONT MATTER ORGANISATION -------------------
\pagenumbering{gobble} % removes page number
\raggedright           % Set the style to left justification, remove for fill justification

\frontmatter
\maketitle
\begin{abstract}
\input{00-abstract}
\end{abstract}
\pdfbookmark[0]{\contentsname}{toc} % Add pdf bookmark for contents page for navigation
\tableofcontents
\listoffigures
\listoftables
\addtotoc{Listings} % The list of listings does not, by default, appear in the ToC
\lstlistoflistings
\listofaddmaterial
\mainmatter
%% ------------------ MAIN MATTER (CONTENT) --------------------
\input{10-introduction}
\input{30-experiment}
\input{50-results}
\input{70-conclusions}

\appendix
\input{90-appendix-a}
\backmatter
\bibliographystyle{plainnat}
\bibliography{references}
\chapter{Bibliography}
To use bibliography as well as the references section use the \texttt{multibbl} package.
\end{document}
%% ----------------------------------------------------------------